---
title: "Assignment 3"
author: "Satyaki Basu Sarbadhikary &Anuroop Roy"
date: "2023-04-27"
output: html_document
---

```{r}
#install.packages('dplyr')
library("dplyr")
library(MASS)
cats
```


```{r}
#extracting heart weight(Hwt)and Body Weight(Bwt)of Female cats
cats=subset(cats,Sex=='F')
cats

#renaming heart wt and body wt of fwmale cats as h and b
h=cats$Hwt
b=cats$Bwt
```
```{r}
calculateStatistic <- function(x,y) {
cor(x,y)
}
x=h
y=b
pseudovalues <- vector(length=length(x), mode='numeric')

stat <- calculateStatistic(x,y)
for (i in 1:length(x)) {
  jack_heart <- x[-i]
  jack_body<-y[-i]
  jackStat <- calculateStatistic(jack_heart,jack_body)
  
  pseudovalues[i] <- stat - (length(x)-1)*(jackStat-stat)
}
#jacknife estimate of the correlation coefficient between body weight and heart weight 
estimate <- mean(pseudovalues)
paste("a.)jacknife estimate of the correlation coefficient between body weight and heart weight= ",estimate)

n <- length(pseudovalues)
#jacknife estimate of Standard error of the correlation coefficient 
SE <- sd(pseudovalues)/sqrt(n)
paste("b.)jacknife estimate of Standard error of the correlation coefficient=",SE)


#bias corrected jacknife estrimate 
bias_jacknife=stat-estimate
paste("c) bias corrected jacknife estimate=",bias_jacknife)









```
Q.7)
 Call the stackloss dataset in R. It gives the operational data of a plant for the
oxidation of ammonia to nitric acid. Information on four variables are given.
Consider the stack.loss as the response variable.


(a) Draw an 80% training data randomly from the dataset. Fit a linear regres-
sion of the response on the predictors. Obtain the predicted values of the 
response using the fitted model on the complementary test set and obtain
the test MSE.
```{r}
rm(list=ls())
set.seed(seed=123)
stackloss
s=stackloss$stack.loss
a=stackloss$Air.Flow
c=stackloss$Acid.Conc.
w=stackloss$Water.Temp

```
```{r}
#drawing 80% training data by random sampling wqithout replacement
sample=sample(c(TRUE,FALSE),nrow(stackloss),replace=T,prob=c(80,30))
train=stackloss[sample,]
test=stackloss[!sample,]
#linear regression of rsponse on predictors
model=lm(stack.loss~Air.Flow+Water.Temp+Acid.Conc.,data=train)
model
test
```
```{r}

#d=data.frame(test$Air.Flow,test$Acid.Conc.,test$Water.Temp)
prediction=predict(model,newdata=test)
#predicted value of response from test data using model of training data
paste("predicted values of response=",prediction)
#test mse
paste("Test Mse=",mean((test$stack.loss-prediction)^2))
```

(b) Repeat the above step 50 times and report the average test MSE.
```{r}
test_mse=array(0)
for(i in 1:50){
  sample=sample(c(TRUE,FALSE),nrow(stackloss),replace=T,prob=c(80,30))
train=stackloss[sample,]
test=stackloss[!sample,]
#linear regression of rsponse on predictors
model=lm(stack.loss~Air.Flow+Water.Temp+Acid.Conc.,data=train)
prediction=predict(model,newdata=test)
#predicted value of response from test data using model of training data
prediction
#test mse
test_mse[i]=mean((test$stack.loss-prediction)^2)
  
}
#
paste("averag test mse=",mean(test_mse))
```
(c) Apply the Leave One Out Cross Validation method to determine the good-
ness of the linear regression of the response on the predictors. Report the
average test MSE.
```{r}
n=dim(stackloss[1])[1]
mse=c()
  for(i in 1:n){
    model1=lm(stack.loss~Air.Flow+Water.Temp+Acid.Conc.,data=stackloss[-i,])
    mse[i]=(stackloss[i,4]-predict(model1,newdata=stackloss[i,]))^2

  }
paste("average test mse=",mean(mse))

```
(d) Randomly divide the dataset of 21 points into k folds. At the i
th stage,consider the data after leaving out the i-th fold as the test set. Fit alinear
regression model on the training set and validate it on the test set and
obtain the test MSE.
(e) Choose k = 5, 7, 10 in the preceeding step and in each case, report the
average test MSE.
```{r}
k5=5
mse5=c()
for(i in 1:k5){
   model5=lm(stack.loss~Air.Flow+Water.Temp+Acid.Conc.,data=stackloss[-i,])
   mse5[i]=(stackloss[i,4]-predict(model1,newdata=stackloss[i,]))^2
}
mean(mse5)

k7=7
mse7=c()
for(i in 1:k7){
   model7=lm(stack.loss~Air.Flow+Water.Temp+Acid.Conc.,data=stackloss[-i,])
   mse7[i]=(stackloss[i,4]-predict(model1,newdata=stackloss[i,]))^2
}
mean(mse7)

k10=10
mse10=c()
for(i in 1:k10){
   model10=lm(stack.loss~Air.Flow+Water.Temp+Acid.Conc.,data=stackloss[-i,])
   mse10[i]=(stackloss[i,4]-predict(model1,newdata=stackloss[i,]))^2
}
mean(mse10)

```
e)Comment on the three methods of cross validation. For the k-fold cross
validation, which value of k yields the least average test MSE?

from the three methods of cross validation,the k fold cross validation approach yields the minimum average test mse for k=10(8.710003)
now the validation set approach yielded largest test mse (24.922)and on repeatedly performing the approach 50 times gave us a relatively smaller test mse(15.483)while leave one out spprosch of cross valiadtion yielded 13.8985 average test mse

