---
title: "Linear Model Cw"
author: "Kanchan Chowdhury"
date: 13.04.2023
output: pdf_document
---

```{r}
setwd("C:/Users/DS-37/Downloads")
d1=read.csv("data1.csv")
d2=read.csv("data2.csv")
d3=read.csv("data3.csv")
```

## Checking Linearity
### Scatterplot
```{r}
plot(d1$X,d1$Y,pch=20)
```
As X and Y are linearly related, we move forward with Linear Regression.
### Fitting Ordinary Learst Square  Model
```{r}
reg_ols1=lm(Y~X,d1)
summary(reg_ols1)
```
### Checking normality of residuals
We apply Shapiro Wilk Test and plot QQ-Plot
```{r}
shapiro.test(reg_ols1$residuals)
qqnorm(reg_ols1$residuals,pch=20)
qqline(reg_ols1$residuals)
```
As, p value = 0.1961 > 0.05. we accept the null hypothesis. Hence we move forward with t-test.
### t-test for $E(\epsilon)=0$
```{r}
t.test(reg_ols1$residuals)
```
Here, the p-value is 1 > 0.05. We strongly accept the null and conclude that mean of the residuals is 0.

### Checking Outliers
```{r}
boxplot(reg_ols1$residuals, data=d1)
```


### Checking Heteroskedasticity
We plot Residual vs Fitted Plot
```{r}
plot(reg_ols1$fitted.values,reg_ols1$residuals,pch=20)
```
### BP Test for OLS model
```{r,warning=FALSE}
library(lmtest)
bgtest(Y ~ X, order=3, data=d1)
```
Here, p-value = 0.002485 < 0.05. Hence, we reject the null hypothesis and conclude that the data is heteroskedastic.
### NCV Test for OLS Model
```{r}
library(car)
ncvTest(reg_ols1)
```
Here, p-value = 0.000018 < 0.05. Hence, we reject the null hypothesis and conclude that the data is heteroskedastic.


Here, the plot is non-random, we move forward with **Weighted Linear Regression**
Here we see that $|e| \propto x$  
V(e) = $\propto x^2$  
WLS -> $w_{ij} = \frac{1}{\sigma_i^2} \propto \frac{1}{x_i^2}$


### Fitting Weighted Learst Square  Model
```{r}
reg_wls1=lm(Y~X,d1,weights = 1/X^2)
summary(reg_wls1)
```
### Plotting Standard residuals vs Fitted Models
```{r}
par(mfrow=c(1,2))
plot(reg_ols1$fitted.values,rstandard(reg_ols1),pch=20,main="OLS Model")
plot(reg_wls1$fitted.values,rstandard(reg_wls1),pch=20,main="WLS Model")
```
### BP Test for WLS Model
```{r}
library(lmtest)
bgtest(reg_wls1, order=3, data=d1)
```
Here, p-value = 0.002485 < 0.05. Hence, we reject the null hypothesis and conclude that the data is heteroskedastic.
### NCV Test for WLS Model
```{r}
ncvTest(reg_wls1)
```
Here, p-value = 0.16617 > 0.05. Hence, we reject the accept hypothesis and conclude that the data is homoskedastic.

## Problem 2
```{r}
reg2=lm(comsales~indsales,d2)
summary(reg2)
```
```{r}
plot(d2$time,reg2$residuals,pch=20)
```

```{r,warning=FALSE}
library(orcutt)
rho=cochrane.orcutt(reg2,convergence=5,max.iter=1000)$rho

y=d2$comsales[-1]-d2$comsales[-20]*rho
x=d2$indsales[-1]-d2$indsales[-20]*rho

reg2=lm(y~x)
```
### Checking autocorrelation
```{r}
#plot(d2$time,reg2$residuals,pch=20)
```
### BG Test
```{r}
library(lmtest)
bgtest(y ~ x, order=3, data=d2)
```
### DW Test
```{r,message=FALSE}
library(car)
durbinWatsonTest(reg2)
```
## Problem 3
```{r}
head(d3)
```
```{r}
d3[,1]
```

```{r}
round(cor(d3[,-c(1:2)]),2)
```
- As Pulse is highly correlated to almost every variable then Pulse should be removed from the model.
```{r}
model=lm(BP~.,data=d3[,-1])
vif(model)
```
- As variance inflation factor is high for weight,BSA and pulse so we should remove these variables from the model.

###Problem 4
Test if the location of sepal length and petal length are same or not.
```{r}
data=iris
head(data)
```
```{r}
library(tidyverse)
setosa=data%>%filter(Species=="setosa")
virginica=data%>%filter(Species=="virginica")
```
#testing for symmetry test.
```{r}
library(lawstat)
symmetry.test(setosa$Sepal.Length-setosa$Petal.Length)
```
```{r}
wilcox.test(setosa$Sepal.Length-setosa$Petal.Length,alternative = "two.sided",paired = F,mu=0)
```
- Diffrence in location exist.
```{r}
g1=ggplot(data=NULL)+
  geom_histogram(aes(x=setosa$Sepal.Length,y=..density..),fill="lightgreen",color="red")+
  geom_histogram(aes(x=setosa$Petal.Length,y=..density..),fill="skyblue",color="black")
  

```



```{r}
symmetry.test(virginica$Sepal.Length-virginica$Petal.Length)
```








